Motion planning for assessing accessibility of protein tunnels

### 2017-09-10 paper.tex
fig 3: graphs on one line
few changes

In this paper, we propose a modification to RRT to address the problem of planning a non-spherical flexible ligand passage through a given protein.
We localize the generation of the random samples, i.e., instead of sampling the whole configuration space we focus only on the space surrounding a potential tunnel detected by the Voronoi-based method (using a smaller ligand probe in order to avoid omitting possibly feasible tunnel candidates).
The potential tunnel is represented by a set of consecutive spheres surrounding the tunnel centerline.
This optimization allows us to find trajectories along the given tunnel, which is faster than searching for any trajectory in the whole protein.
To enable the movement of the ligand in the narrow parts of the tunnel, the free-space can be dilated by shrinking the atom radii, similarly, e.g., to~\cite{cortes2005path,hsu06multilevel}. 
The ligand flexibility is modeled using a fixed set of predefined conformations which the ligand can occupy.

Our proposed method was tested by the biochemists who compared the results with trajectories of ligands determined experimentally in the lab.

@@

\section{Proposed Method}
To address the problems of the existing solutions for computing the non-spherical flexible ligand trajectories through protein tunnels, we propose a modification of the RRT planner. 
Our approach takes into account the shape and flexibility of the ligand and localizes the searched configuration space using the information about tunnels detected by the Voronoi-based approach.
The input is therefore formed by a set of tunnels computed using Voronoi diagrams and the task is to find trajectories for a ligand along these tunnels
leading from outside the protein to the active site.
The tunnels are detected in a sequence time steps capturing protein dynamic behavior that is computed using molecular dynamics (MD) simulations.

@@

\begin{figure}[bh]
\centering
%\includegraphics[width=0.25\textwidth]{fig/m040-conformations}
\includegraphics[width=0.14\textwidth]{fig/m040-conf1} \hskip 25pt
\includegraphics[width=0.11\textwidth]{fig/m040-conf2}
\caption{\label{fig::m040c}
 The examples of a `long' conformation (left) and a `packed' one for the ligand m040 (1,5-Dichloropentane).
}
\end{figure}

@@

\begin{table}[ht]
\centering
\caption{\label{tab::one}Traversability for one atom.}
\renewcommand{\tabcolsep}{4.3pt}
{\small
\input graphProbe.tex
}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{\label{tab::selection}
    The influence of the conformation selection method to the traversability rates in the first tunnel.
    The number after~`$/$' denotes the number of atoms.
}
\small
\renewcommand{\tabcolsep}{4.3pt}
{\small
\input graphSel.tex
}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{\label{tab::main}
    Traversability of the tunnels for ligands with 50 conformations. 
    The number after '$/$' denotes the number of atoms.
}
\small
\renewcommand{\tabcolsep}{4.3pt}
{\small
\input graph.tex
}
\end{table}

@@

{\def\a{270}
\begin{figure}[hb]
\centering
\includegraphics[width=0.25\textwidth,angle=\a]{fig/m037tmins-03tunnel-1rrt_path_time-2}
\includegraphics[width=0.25\textwidth,angle=\a]{fig/m037tmins-03tunnel-1trajDistToGoal-2}
\caption{\label{fig::comparison}
    Runtimes and the distance to the active site for the m037 ligand at scale 0.6.
    The values are computed from all 200 trials (20 initial configurations $\times$ 10 trials).
    Only each second frame is shown in the graph.
}
\end{figure}
}

>>>>

In this paper, we propose a modification to RRT to address the problem of planning a non-spherical flexible ligand passage through a given protein.
We localize the generation of the random samples, i.e., instead of sampling the whole configuration space we focus only on the space surrounding a potential tunnel detected by the Voronoi-based method (using a smaller ligand probe in order to avoid omitting possibly feasible tunnel candidates).
The potential tunnel is represented by a set of consecutive spheres surrounding the tunnel centerline.
This optimization allows us to find trajectories along the given tunnel, which is faster than searching for any trajectory in the whole protein.
To enable the movement of the ligand in the narrow parts of the tunnel, the free-space can be dilated by shrinking the atom radii, similarly, e.g., to~\cite{cortes2005path,hsu06multilevel}. 
The ligand flexibility is modeled using a fixed set of predefined conformations which the ligand can occupy.
Our proposed method was tested by the biochemists who compared the results with trajectories of ligands determined experimentally in the lab.

@@

\section{Proposed Method}
To address the problems of the existing solutions for computing the non-spherical flexible ligand trajectories through protein tunnels, we propose a modification of the RRT planner. 
Our approach takes into account the shape and flexibility of the ligand and localizes the searched configuration space using the information about tunnels detected by the Voronoi-based approach.
The input is therefore formed by a set of tunnels computed using Voronoi diagrams and the task is to find trajectories for a ligand along these tunnels
leading from outside the protein to the active site.
The tunnels are detected in a sequence time steps capturing protein dynamic behavior computed using MD simulations.

@@

\begin{figure}[b]
\centering
%\includegraphics[width=0.25\textwidth]{fig/m040-conformations}
\includegraphics[width=0.14\textwidth]{fig/m040-conf1} \hskip 25pt
\includegraphics[width=0.11\textwidth]{fig/m040-conf2}
\caption{\label{fig::m040c}
 The examples of a `long' conformation (left) and a `packed' one for the ligand m040 (1,5-Dichloropentane).
}
\end{figure}

@@

\begin{table}[ht]
\centering
\caption{\label{tab::one}Traversability for one atom.}
\renewcommand{\tabcolsep}{4.3pt}
{\small
\renewcommand{\arraystretch}{0.7}
\input graphProbe.tex
}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{\label{tab::selection}
    The influence of the conformation selection method to the traversability rates in the first tunnel.
    The number after~`$/$' denotes the number of atoms.
}
\small
\renewcommand{\tabcolsep}{4.3pt}
{\small
\renewcommand{\arraystretch}{0.7}
\input graphSel.tex
}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{\label{tab::main}
    Traversability of the tunnels for ligands with 50 conformations. 
    The number after '$/$' denotes the number of atoms.
}
\small
\renewcommand{\tabcolsep}{4.3pt}
{\small
\renewcommand{\arraystretch}{0.7}
\input graph.tex
}
\end{table}

@@

{\def\a{270}
\begin{figure}[hb]
\centering
%\includegraphics[width=0.17\textwidth,angle=\a]{fig/m037tmins-03tunnel-1rrt_path_time-2}
%\includegraphics[width=0.17\textwidth,angle=\a]{fig/m037tmins-03tunnel-1trajDistToGoal-2}
\includegraphics[width=0.18\textwidth,angle=\a]{fig/crop1}
\includegraphics[width=0.18\textwidth,angle=\a]{fig/crop2}
\caption{\label{fig::comparison}
    Runtimes and the distance to the active site for the m037 ligand at scale 0.6.
    The values are computed from all 200 trials (20 initial configurations $\times$ 10 trials).
    Only each second frame is shown in the graph.
}
\end{figure}
}
